# Data

## Description

Please note that in this project, we use the term movie loosely. Although we pulled from a movie dataset, many of the rows of data existed as TV shows or Mini Series. We utilize the term movies to encompass the idea of film production due to the nature of the data and for the sake of simplicity.

The data we are using primarily are "MovieLens" data. They were collected by GroupLens Research, a research group at the University of Minnesota, Twin Cities, aimed to provide recommendations of movies for users. The dataset we are using is called MovieLens 32M, which comprises of 32 million ratings and 2 million tags applied to 87,585 titles by 200,948 users. It includes “links.csv,” which assigns each movie an ID, and also lists out each movie's corresponding ID on IMDb and TMDB (both are movie-rating websites). “movies.csv” comprises of each movie’s title and genres.  “ratings.csv” documents each user’s rating of the movies they have seen. “tags.csv” comprises of users' tags to the movies, expressed in words or short phrases. The dataset was collected in October 2023. We downloaded it directly from GroupLens' website. In addition, we are going to cross-reference data from IMDB. We utilized the title.basics.tsv dataset located on the IMDB developer website. This dataset helped us get more information on our existing data. This dataset contains the variables tconst which is a unique identifier for the variable (we used this to help join datasets) and titleType which was the format of type of the movie production. There were additional title variables like primaryTitle which was the most popular title and original title which was the original title of the movie. There was also distinguishing characteristics of adult titles and non-adult titles as well as the starting year and ending year for releases. Finally, a column called runtimeMinutes tracks the runtime of the production. 

## Missing value analysis

```{r}
variable='/Users/erinma/Documents/STAT5702/ml-32m/'
movies=read.csv('/Users/erinma/Documents/STAT5702/ml-32m/movies.csv')
ratings=read.csv('/Users/erinma/Documents/STAT5702/ml-32m/ratings.csv')
```

```{r}
result <- merge(movies, ratings, by = "movieId", all = TRUE)
```

```{r}
summary(result)

```

```{r}
write.csv(result, "result.csv")
```


```{r}
library(tidyverse)

na_counts <- colSums(is.na(result))

na_df <- data.frame(
  Column = names(na_counts),
  NA_Count = na_counts
)

ggplot(na_df, aes(x = Column, y = NA_Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Number of NA Values by Column",
    x = "Column",
    y = "NA Count"
  )
```

After merging the dataset movies with the dataset ratings, the graph and summary plot show us that there are around 3000 NA values in rating, timestamp, and userId. These are all columns in the original ratings dataset. This means that there are many movies that do not have ratings by any users. 

```{r}
links <- read.csv(file.path(variable, 'links.csv'))

na_counts <- data.frame(
  Column = c("imdbId", "tmdbId"),
  NA_Count = c(sum(is.na(links$imdbId)), sum(is.na(links$tmdbId)))
)

ggplot(na_counts, aes(x = Column, y = NA_Count, fill = Column)) +
  geom_bar(stat = "identity", color = "black") +
  labs(
    title = "Count of NA Values in imdbId and tmdbId",
    x = "Column",
    y = "Number of NA Values"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")
```
The "links.csv" file shows each movie's corresponding ID in IMDb and TMDB. The graph above shows that some movies do not have corresponding IDs in TMDB, because those movies are not found on TMDB.
