---
title: "Exploratory Analysis"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}
movies <- read.csv("/Users/melodyma/Downloads/ml-32m/movies.csv")
all_genres <- unlist(strsplit(as.character(movies$genres), "\\|"))
unique_genres <- unique(all_genres)
print(unique_genres)
```
```{r}
movies <- movies[movies$genres != "(no genres listed)",]
all_genres <- unlist(strsplit(as.character(movies$genres), "\\|"))
unique_genres <- unique(all_genres)
print(unique_genres)
```

```{r}
genre_list <- strsplit(movies$genres, "\\|")
movies_ohe <- movies
for (genre in unique_genres) {
  movies_ohe[[genre]] <- sapply(genre_list, function(x) as.integer(genre %in% x))
}
```

```{r}
movies_ohe <- movies_ohe[, !names(movies_ohe) %in% "imax"]
```


```{r}
ratings <- read.csv("/Users/melodyma/Downloads/ml-32m/ratings.csv")
```

```{r}
colnames(user_movie_data)
```


```{r}
# Step 1: Merge `ratings` and `movies_ohe` on `movieId`
user_movie_data <- merge(ratings, movies_ohe, by = "movieId")

# Step 2: Group by `userId` and calculate the sum for each genre and the total movies rated
library(dplyr)
user_genre_proportions <- user_movie_data |> 
  group_by(userId) |> 
  summarise(
    Total_Movies = n(), # Total number of movies rated by each user
    across(Adventure:`Film-Noir`, sum, na.rm = TRUE) # Sum of each genre
  ) |> 
  mutate(
    across(Adventure:`Film-Noir`, ~ .x / Total_Movies) # Proportion of each genre
  )

# Step 3: Remove the intermediate "Total_Movies" column (optional)
user_genre_proportions <- user_genre_proportions |> 
  select(-Total_Movies)

write.csv(user_genre_proportions, "~/EDAV Movies/Exploratory Work/user_genre_proportions.csv", row.names = FALSE)


# Print the resulting dataframe
print(user_genre_proportions)

```


```{r}
# Example: Assuming your data frame is named user_genre_proportions
# Exclude the userId column
data_corr <- user_genre_proportions[, !names(user_genre_proportions) %in%  "userId"]

# Get all pairs of column names
column_pairs <- combn(names(data_corr), 2, simplify = FALSE)

# Initialize an empty list to store correlation results
correlations <- list()

# Loop through each pair of columns and calculate the correlation
for(pair in column_pairs) {
  correlation <- cor(data_corr[[pair[1]]], data_corr[[pair[2]]], use = "complete.obs")
  correlations[[paste(pair[1], pair[2], sep = "_")]] <- correlation
}

# Print correlations
print(correlations)
```


```{r}
install.packages("gridExtra")
library(gridExtra)
```

```{r}
library(ggplot2)
```


```{r}

data <- user_genre_proportions[, !names(user_genre_proportions) %in% "userId"]

# Make sure all column names are valid (e.g., no special characters)
colnames(data) <- make.names(colnames(data))

# Select the columns for the specific pairs you want to plot
pairs_to_plot <- list(
  c("Adventure", "Fantasy"),
  c("Adventure", "Action"),
  c("Animation", "Children"),
  c("Animation", "Fantasy"),
  c("Children", "Fantasy"),
  c("Children", "Musical"),
  c("Action", "Sci.Fi")  # Ensure the name matches after make.names() transformation
)

# Create individual scatter plots for each pair
plot_list <- lapply(pairs_to_plot, function(pair) {
  ggplot(data, aes(x = .data[[pair[1]]], y = .data[[pair[2]]])) +
    geom_point() +
    labs(title = paste(pair[1], "vs", pair[2])) +
    theme_minimal()
})

# Arrange the plots into one panel
grid.arrange(grobs = plot_list, ncol = 3)

```

```{r}
install.packages("grid")
library("grid")
```


```{r}
data <- user_genre_proportions[, !names(user_genre_proportions) %in% "userId"]

# Ensure the column names are valid (e.g., no special characters)
colnames(data) <- make.names(colnames(data))

# Select the columns for the specific pairs you want to plot (removing "Children" and "Fantasy")
pairs_to_plot <- list(
  c("Adventure", "Fantasy"),
  c("Adventure", "Action"),
  c("Animation", "Children"),
  c("Animation", "Fantasy"),
  c("Children", "Musical"),
  c("Action", "Sci.Fi")  # Keep Sci.Fi as is since make.names() converts it
)

# Calculate correlations for each pair
correlations <- sapply(pairs_to_plot, function(pair) {
  cor(data[[pair[1]]], data[[pair[2]]], use = "complete.obs")
})

# Create individual scatter plots for each pair, adding the correlation value as text
plot_list <- lapply(1:length(pairs_to_plot), function(i) {
  pair <- pairs_to_plot[[i]]
  
  # Replace Sci.Fi with Sci-Fi in the titles
  plot_title <- gsub("Sci.Fi", "Sci-Fi", paste(pair[1], "vs", pair[2]))
  
  ggplot(data, aes(x = .data[[pair[1]]], y = .data[[pair[2]]])) +
    geom_point() +
    labs(title = plot_title) +  # Only add the title, leave axis labels unchanged
    annotate("text", x = max(data[[pair[1]]]), y = max(data[[pair[2]]]), 
             label = paste("r =", round(correlations[i], 2)), 
             hjust = 1, vjust = 1, color = "blue") +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 9),  # Smaller plot title
      axis.title.x = element_text(size = 8),  # Smaller x-axis title
      axis.title.y = element_text(size = 8),  # Smaller y-axis title
      plot.margin = margin(t = 15, r = 5, b = 5, l = 5)  # Increase top margin
    )
})

# Arrange the plots into one panel with adjusted margins
grid.arrange(grobs = plot_list, ncol = 3)

# Adjust the position of the panel title to avoid clipping
grid.text("Correlation between Genre Selection", 
          x = 0.5, y = 0.982, gp = gpar(fontsize = 13, fontface = "bold"))
```


```{r}
# Assuming user_genre_proportions is the dataframe with genre proportions
# Exclude the userId column to focus on the genre proportions
genre_proportions <- user_genre_proportions |> 
  select(-userId)

# Get the five-number summary for all genre proportions
five_number_summary <- apply(genre_proportions, 2, fivenum)

# Print the five-number summary
print(five_number_summary)
```


```{r}
library(GGally)
library(ggplot2)
```


```{r}
user_genre_proportions$userId <- as.factor(user_genre_proportions$userId)
ggparcoord(
  data = user_genre_proportions,
  columns = 2:ncol(user_genre_proportions),  # Select genre columns
  alphaLines = 0.2,
  splineFactor = TRUE
) +
  ggtitle("Parallel Coordinate Plot of User Genre Proportions") +
  xlab("Genres") +
  ylab("Proportions") +
  theme_minimal() +
  scale_color_manual(values = "grey")
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


```{r}
# List of genres (common across user_movie_data)
genres <- c("Adventure", "Animation", "Children", "Comedy", "Fantasy", "Romance", 
            "Drama", "Action", "Crime", "Thriller", "Horror", "Mystery", "Sci-Fi", 
            "IMAX", "Documentary", "War", "Musical", "Western", "Film-Noir")

# Function to calculate average rating for a specific genre
calculate_genre_avg_rating <- function(data, genre) {
  # Filter for rows where the movie belongs to the specified genre
  genre_data <- data[data[[genre]] == 1, ]
  
  # Calculate average rating for each user
  avg_ratings <- aggregate(rating ~ userId, data = genre_data, mean)
  
  # Rename the column for clarity
  colnames(avg_ratings)[2] <- paste0("avg_rating_", genre)
  
  return(avg_ratings)
}

# Calculate average ratings for all genres and combine results
all_genre_avg_ratings <- Reduce(function(x, y) merge(x, y, by = "userId", all = TRUE), 
                                 lapply(genres, function(genre) {
                                   calculate_genre_avg_rating(user_movie_data, genre)
                                 }))

# View the resulting dataframe
head(all_genre_avg_ratings)

```


```{r}
library(ggplot2)
library(reshape2)

# Step 1: Merge the dataframes on 'userId'
merged_data <- merge(all_genre_avg_ratings, user_genre_proportions, by = "userId")

# Step 2: Reshape the data to long format for plotting
# Reshape average ratings
avg_ratings_long <- melt(merged_data, 
                         id.vars = "userId", 
                         measure.vars = paste0("avg_rating_", genres), 
                         variable.name = "genre", 
                         value.name = "avg_rating")

# Reshape genre proportions
genre_proportions_long <- melt(merged_data, 
                               id.vars = "userId", 
                               measure.vars = genres, 
                               variable.name = "genre", 
                               value.name = "proportion")

# Step 3: Combine the reshaped data
# Standardize genre column names to match (remove "avg_rating_" prefix in avg_ratings_long)
avg_ratings_long$genre <- gsub("avg_rating_", "", avg_ratings_long$genre)
combined_data <- merge(avg_ratings_long, genre_proportions_long, by = c("userId", "genre"))

# Step 4: Create the scatterplot
ggplot(combined_data, aes(x = proportion, y = avg_rating, color = genre)) +
  geom_point(alpha = 0.7) +
  labs(title = "Scatterplot of User Proportions vs. Average Ratings by Genre",
       x = "Proportion of Movies Watched (Genre)",
       y = "Average Rating (Genre)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_discrete(name = "Genre")
```


```{r}
# Step 1: Merge the dataframes on 'userId'
merged_data <- merge(all_genre_avg_ratings, user_genre_proportions, by = "userId")

# Step 2: Reshape the data to long format for plotting
# Reshape average ratings
avg_ratings_long <- melt(merged_data, 
                         id.vars = "userId", 
                         measure.vars = paste0("avg_rating_", genres), 
                         variable.name = "genre", 
                         value.name = "avg_rating")

# Reshape genre proportions
genre_proportions_long <- melt(merged_data, 
                               id.vars = "userId", 
                               measure.vars = genres, 
                               variable.name = "genre", 
                               value.name = "proportion")

# Step 3: Combine the reshaped data
# Standardize genre column names to match (remove "avg_rating_" prefix in avg_ratings_long)
avg_ratings_long$genre <- gsub("avg_rating_", "", avg_ratings_long$genre)
combined_data <- merge(avg_ratings_long, genre_proportions_long, by = c("userId", "genre"))

# Step 4: Filter data to include only rows where the proportion is greater than 0.7
filtered_data <- combined_data %>%
  filter(proportion > 0.7)

# Step 5: Create the scatterplot
ggplot(filtered_data, aes(x = proportion, y = avg_rating, color = genre)) +
  geom_point(alpha = 0.7) +
  labs(title = "Scatterplot of User Proportions vs. Average Ratings by Genre",
       x = "Proportion of Movies Watched (Genre)",
       y = "Average Rating (Genre)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_discrete(name = "Genre")
```


```{r}
# Step 1: Merge the dataframes on 'userId'
merged_data <- merge(all_genre_avg_ratings, user_genre_proportions, by = "userId")
file_path <- "~/EDAV Movies/Exploratory Work/Prop_AvgRating.csv"
write.csv(merged_data, file = file_path, row.names=FALSE)
merged_data <- read.csv("~/EDAV Movies/Exploratory Work/Prop_AvgRating.csv")
colnames(merged_data)
```



```{r}

# Step 2: Reshape the data to long format for plotting
# Reshape average ratings
avg_ratings_long <- melt(merged_data, 
                         id.vars = "userId", 
                         measure.vars = paste0("avg_rating_", genres), 
                         variable.name = "genre", 
                         value.name = "avg_rating")

# Reshape genre proportions
genre_proportions_long <- melt(merged_data, 
                               id.vars = "userId", 
                               measure.vars = genres, 
                               variable.name = "genre", 
                               value.name = "proportion")

# Step 3: Combine the reshaped data
# Standardize genre column names to match (remove "avg_rating_" prefix in avg_ratings_long)
avg_ratings_long$genre <- gsub("avg_rating_", "", avg_ratings_long$genre)
combined_data <- merge(avg_ratings_long, genre_proportions_long, by = c("userId", "genre"))

# Step 4: Create a binary indicator for proportion >= 0.7
combined_data$proportion_group <- ifelse(combined_data$proportion >= 0.7, "Genre Proportion >= 0.7", "Genre Proportion < 0.7")

# Step 5: Create the boxplot
ggplot(combined_data, aes(x = proportion_group, y = avg_rating, fill = proportion_group)) +
  geom_boxplot() +
  labs(title = "Average Rating Distribution and Genre Preferences",
       x = "Proportion Group",
       y = "Average Rating") +
  theme_minimal() +
  scale_fill_manual(values = c("Genre Proportion >= 0.7" = "skyblue", "Genre Proportion < 0.7" = "salmon")) +
  theme(legend.position = "none")
```


```{r}
install.packages("tidytext")
install.packages("syuzhet")
```


```{r}
library(tidytext)
library(dplyr)
library(syuzhet)
```


```{r}
tags <- read.csv("/Users/melodyma/Downloads/ml-32m/tags.csv")
tags_sentiment <- tags |> 
  mutate(sentiment_score = get_sentiment(tag, method = "afinn"))

print(tags_sentiment)
```


```{r}
tags_sentiment_syuzhet <- tags |> 
  mutate(sentiment_score = get_sentiment(tag, method = "syuzhet"))

print(tags_sentiment_syuzhet)
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

